Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-02-21 paper/ms.tex
a few more words about astero example [ci skip]

% Projects:
\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\celerite}{\project{celerite}}
\newcommand{\emcee}{\project{emcee}}

@@

In this first example, we will simulate a dataset using a known \celerite\
process and fit it with \celerite\ to demonstrate that valid inferences can be
made in this idealized case.

@@

@@

To demonstrate this method, we will use a very simple heuristic model based on
\dfmtodo{CITE} where the model PSD is given by a mixture of 6 components with
amplitudes and frequencies specified by $\nu_\mathrm{max}$, $\Delta \nu$, and
several nuisance parameters.
The first term is used to capture the granulation ``background'' using
\eq{granulation-psd} with two free parameters $S_g$ and $\omega_g$.
The remaining 5 terms are given by \eq{sho-psd} where $Q$ is a nuisance
parameter shared between terms and the frequencies are given by
\begin{eqnarray}
\omega_{0,\,j} = 2\,\pi\,(\nu_\mathrm{max} + j\,\Delta\nu + \epsilon)
\end{eqnarray}
and the amplitudes are given by
\begin{eqnarray}
S_{0,\,j} =
    \frac{A}{Q^2}\,\exp\left(-\frac{[j\,\Delta\nu + \epsilon]^2}{2\,W^2}\right)
\end{eqnarray}
where $j$ is an integer running from $-3$ to 3 and $\epsilon$, $A$, and $W$
are shared nuisance parameters.
This model could be easily extended to include small frequency splitting and
$\nu_\mathrm{max}$ and $\Delta \nu$ could be replaced by physical parameters
like $\log g$.

KIC~11615890 \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz}
\end{eqnarray}

\dfmtodo{Add plots etc.}

>>>>

% Projects:
\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\celerite}{\project{celerite}}
\newcommand{\emcee}{\project{emcee}}

@@

In this first example, we will simulate a dataset using a known \celerite\
process and fit it with \celerite\ to demonstrate that valid inferences can be
made in this idealized case.
The simulated dataset is shown in the left panel of \Figure{simulated-correct}
and it was generated using a SHO kernel (\eqalt{sho-kernel}) with parameters
$S_0 = 1$, $\omega_0 = e^2$, and $Q = e^2$.
The true PSD is shown as a dashed line in the right panel of
\Figure{simulated-correct}.
We applied log-uniform priors to all of the parameters and used \emcee\
\citep{Foreman-Mackey:2013} to sample the joint posterior density and computed
the marginalized posterior inference of the PSD.
This inference is shown in the right panel of \Figure{simulated-correct} as a
blue contour indicating 68\% of the posterior mass.
It is clear from this figure that, as expected, the inference correctly
reproduces the true PSD.

@@

In this section we will demonstrate several use cases of \celerite\ when
applied to real datasets.
Each of these examples touches on an active area of research so we will limit
our examples to be qualitative in nature and make no claim of optimality but
we hope that these examples will encourage interested readers to investigate
the applicability of \celerite\ to their research.

All of the following examples show time domain datasets with a clear bias in
favor of large homogeneous photometric surveys but these methods can similarly
be applied to spectroscopy, where wavelength~--~instead of time~--~is the
independent coordinate and other one-dimensional domains.
It is also possible to cast some two-dimensional problems in the \celerite\
framework (\dfmtodo{CITE Ian C}).

@@

To demonstrate this method, we will use a very simple heuristic model based on
\dfmtodo{CITE} where the model PSD is given by a mixture of 8 components with
amplitudes and frequencies specified by $\nu_\mathrm{max}$, $\Delta \nu$, and
several nuisance parameters.
The first term is used to capture the granulation ``background'' using
\eq{granulation-psd} with two free parameters $S_g$ and $\omega_g$.
The remaining  7 terms are given by \eq{sho-psd} where $Q$ is a nuisance
parameter shared between terms and the frequencies are given by
\begin{eqnarray}
\omega_{0,\,j} = 2\,\pi\,(\nu_\mathrm{max} + j\,\Delta\nu + \epsilon)
\end{eqnarray}
and the amplitudes are given by
\begin{eqnarray}
S_{0,\,j} =
    \frac{A}{Q^2}\,\exp\left(-\frac{[j\,\Delta\nu + \epsilon]^2}{2\,W^2}\right)
\end{eqnarray}
where $j$ is an integer running from $-3$ to 3 and $\epsilon$, $A$, and $W$
are shared nuisance parameters.
This model could be easily extended to include small frequency splitting and
$\nu_\mathrm{max}$ and $\Delta \nu$ could be replaced by physical parameters
like $\log g$.

To demonstrate the applicability of this model, we apply it to infer the
asteroseismic parameters of the giant star KIC~11615890 observed by the
\kepler\ Mission.
The goal of this example is to show that, even for a low signal-to-noise
dataset with a short baseline, it is possible to infer asteroseismic
parameters with formal uncertainties that are consistent with the parameters
inferred with a much larger dataset.
Looking forward to \tess, we will infer $\nu_\mathrm{max}$ and $\Delta\nu$
using only one month of \kepler\ data and compare our results to the results
inferred from the full 4 year baseline of the mission.
For KIC~11615890, the published asteroseismic parameters based on the full
dataset are \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz} \quad.
\end{eqnarray}
We randomly select a month-long segment of \kepler\ data, initialize our
\celerite\ model using a grid search in the parameter space, and then use
\emcee\ \citep{Foreman-Mackey:2013} to sample the joint posterior density for
the full set of parameters.
\Figure{astero-corner} shows the marginalized density for $\nu_\mathrm{max}$
and $\Delta\nu$ compared to the results from the literature.
\dfmtodo{Add some more discussion here.}
